Agile Monitoring of Evolving Web Services: For the Little Man

### 2018-04-05 architecture.tex
smarter table
cleaning up...

@@

  % This used to be in the first section... 
  % Maybe it fits better localized here: 
  Data collected by the wrappers are persisted in a local database. The SQLAlchemy Object Relational Mapper\footnote{\url{https://www.sqlalchemy.org/}} on top of SQLite\footnote{\url{https://www.sqlite.org/}} is used for this purpose.

>>>>

The viewpoints we showed are only a subset of all the viewpoints of the \tool. In practice, the tool provides other perspectives as long as they are a combination of:

\begin{itemize}
  \item Cardinality: one endpoint vs multiple
  \item Objective: utilization vs. performance (response time)
  \item Evolution Axis: time vs. versions
  \item Grouping: Grouped (e.g. by user) vs. All
\end{itemize}

@@

  % This used to be in the first section... 
  % Maybe it fits better localized here: 
  Data collected by the wrappers are persisted in a local database. The SQLAlchemy Object Relational Mapper\footnote{\url{https://www.sqlalchemy.org/}} allows for a DB system independent solution.

  By default the system is deployed with an SQLite database \footnote{\url{https://www.sqlite.org/}} is used for this purpose.

### 2018-04-05 introductory.tex
cleaning up...

\begin{abstract}
\todo{Update abstract \& title}
  Python is one of the fastest growing programming languages of the moment. Flask, a Python-based web framework is the technology chosen by tens of thousands of web applications. Many of these Flask-based projects are small websites which can not afford to deploy an advanced service monitoring solution because they are too complicated to install.

  In this paper, investigate how simple can it be to start monitoring a Flask API by developing and presenting \tool, a library that aims to provide insight into the utilization and performance of evolving services. We present the ease with which the library can be integrated in an already existing web application, discuss some of the visualization perspectives that the library provides and point to some future challenges for similar libraries.

>>>>

\begin{abstract}
  Python is one of the fastest growing programming languages of the moment. Flask, a Python-based web framework is the technology chosen by tens of thousands of web applications. Many of these Flask-based projects are small websites which can not afford to deploy an advanced service monitoring solution because they are too complicated to install. Many of these projects might offer a service API. Given the complexity of professional monitoring tools it is very likely that

  In this paper, we set out to show how to design a dashboard for a Flask API which is useful in understanding and guiding service evolution while at the same time, requiring minimal configuration and installation. We present \tool, a tool that provide insight into the utilization and performance of evolving services. We present the ease with which the library can be integrated in an already existing web application, discuss some of the visualization perspectives that the library provides and point to some future challenges for similar libraries.

### 2018-04-05 loc1.tex
done or the day
cleaned up the performance testing part
cleaning up...

%!TEX root=restructured.tex

  \newpage
  \section{LOC \#1: Service Utilization and Performance}

  Since it is taylored for Flask applications, to deploy the \tool one simply needs to use one line of code\footnote{We don't count the import statement that enables that line...} to bind the dashboard to their Flask web application\footnote{\ins{ In this paper we present the integration with APIs written in Python and Flask hoping that this will not prevent the reader from seeing the more general idea; all the tools we show here for Flask can be applied to other API technologies (e.g. Django) by simply providing a few back-end adapters in the right places. }}:

@@

  # app is main application object
  # every Flask web application has one
  dashboard.bind(app)

@@

  During binding, the \tool will search for all endpoints defined in the target application 
    \ins{ 
      and add function wrappers around all the corresponding endpoint functions
    }
% 
  % In order to monitor an endpoint, the \tool creates a function wrapper for the API function that corresponds to the endpoint. This way, the wrapper will be executed whenever that API call is made before the actual function is called. The wrapper contains the code that takes care of monitoring an endpoint. 
%

   \ins{
    The tool takes advantage of the fact that the monitored API already has a web presence, and makes available one extra endpoint (i.e. \code{/dashboard}), which serves the interactive visualization perspectives presented in the remainder of this paper. 
  }  
 The first perspective presents all the automatically discovered endpoints and lets the user select the ones that should be monitored. 
  \Fref{fig:sep} shows that the last access time of every endpoint is tracked irrespective of whether it is selected by the user to be monitored or not.

@@

  \ins{
    We have decided for an opt-in approach to monitoring to avoid any the performance penalties incurred by the dashboard to affect performance sensitive endpoints. We discuss performance issues later. Also, we discuss later one situation in which opt-out might be desirable. 
  }

  \ins{

    One alternative to allowing the user to use annotations would be to let them to annotate the code. However, this pollutes the code, and prevents deploying two versions which would monitor different endpoints.

  }

@@

  \ins
  {

      The remainder of this section presents several of the interactive
      visualizations that become available without any further configuration\footnote{We recommend obtaining a color version of this paper for better readability}.

  }

>>>>

%!TEX root=paper.tex

  \newpage
  \section{Service Utilization and Performance}

  % Since it is tailored for Flask applications, the deployment of \tool is trivial.

  The tool is designed to take advantage of the fact that it runs in the same web application context and that the monitored API already has a web presence. It thus simply makes available one extra endpoint (i.e. \code{/dashboard}) at the same address as the API, endpoint which serves the interactive application presented in the remainder of this paper.

  A dashboard can be attached to an already existing Flask application with a single line of code as the code in the following snippet shows
  \footnote{We do not count the import statement that enables that line}:

  % which binds the dashboard to a Flask application\footnote{\ins{ In this paper we present the integration with APIs written in Python and Flask hoping that this will not prevent the reader from seeing the more general idea; all the tools we show here for Flask can be applied to other API technologies (e.g. Django) by simply providing a few back-end adapters in the right places. }}:

@@

  # LOC #1: associate the main Flask application 
  # object with the dashboard
  dashboard.bind(app)

@@

  During binding, the \tool will search for all endpoints defined in the target application and make them available in the dashboard in an interactive configuration panel (\Fref{fig:sep}). The panel presents all the automatically discovered endpoints and lets the user select the ones that should be monitored (the checkboxes in the last column). When the user enables the monitoring of an endpoint a function wrapper is added to the implementation of that endpoint (which is a function in Flask). The information is also saved in a database such that it is persistent across application restarts.

  % In order to monitor an endpoint, the \tool creates a function wrapper for the API function that corresponds to the endpoint. This way, the wrapper will be executed whenever that API call is made before the actual function is called. The wrapper contains the code that takes care of monitoring an endpoint.

@@

  By default the dashboard takes an {\em opt-in} approach to monitoring: to prevent the performance penalties incurred by the dashboard\footnote{We discuss performance issues later} to affect performance sensitive endpoints, the service developer is asked to manually add the endpoints they want to monitor.

  One alternative to configuring the monitored endpoints from the user interface is to allow the service developer to annotate the code. This would however pollute the code, and prevent deploying two versions which would monitor different endpoints. Also, would pose problems with endpoint evolution in the future.

@@

  The remainder of this section presents several interactive
  visualizations that are available in the dashboared without any further configuration\footnote{We recommend obtaining a color version of this paper for better readability}.
